---
title: "Assignment 5: Substance Use Treatment Mapping"
subtitle: "Patrick Cheng"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
toc: true
---

## Overview

Due: **Friday 10/2, 11:59pm**

In this assignment, you will apply analytic and mapping skills to the 2024 TEDS-D dataset from SAMHSA.

Refer to the following files for reference, though some of what is asked of you here will require logic beyond what we discussed in class:

- week06-treatment-mapping-part1.qmd

- week06-treatment-mapping-part2.qmd

**Before you start:** keep the [codebook](https://www.samhsa.gov/data/system/files/media-puf-file/TEDS-D-2024-DS0001-info-codebook_v1_0.pdf) open in another tab. Every question below depends on knowing what the values in a column actually mean. Don't assume — look it up.

**Time estimate:** about three-four hours total. Question 1 should take you ten minutes. If you are stuck on one thing for more than twenty minutes, ask your classmates or get in touch with me. Being stuck is normal. Staying stuck quietly by yourself is not :)

## Load Packages and Data

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(sf) # "simple features" for spatial data
library(urbnmapr) # access to US state/country shapefiles for mapping
library(ggiraph) # makes ggplot charts interactive
options(scipen = 99) # turns off scientific notation
```

This is the cleaned dataset you saved at the end of Part 1.

```{r}
#| label: read-data

teds_d <- readRDS("data/teds_d_clean.rds")
```

## Question 1: Warm-up

Before mapping anything, get oriented in the data.

a)  Using `tabyl()`, show the breakdown of `reason` (reason for discharge). Which reason is most common, and what percentage of all discharges does it account for?

b)  Make one bar chart showing the number of discharges by `services_d` (type of treatment service). Give it an informative title and axis labels.

c)  In a sentence or two: what does the `reason` value "Transferred to another treatment program or facility" mean for our completion percentages? If a state transfers a lot of patients, would its completion rate look higher or lower than a state that transfers very few?

> **Full credit:** a working `tabyl()`, a readable labeled bar chart, and a sentence or two showing you thought about part (c). There is no trick in this question!

```{r}
#| label: q1.a

### Using tabyl(), show the breakdown of reason (reason for discharge). Which reason is most common, and what percentage of all discharges does it account for?
class(teds_d$reason)
#look at this variable; check class
summary(teds_d$reason)
#summary of data frame

teds_d <- teds_d %>%
  mutate(reason = recode(as.character(reason),
                         "1" = "Treatment completed",
                         "2" = "Dropped out of treatment",
                         "3" = "Terminated by facility",
                         "4" = "Transferred to another treatment program or facility",
                         "5" = "Incarcerated",
                         "6" = "Death",
                         "7" = "Other",
                         .missing = "unknown"),
         reason = factor(reason))
# ////////////////////////////////////////////////////////////////////////////
# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

teds_d %>% 
  tabyl(reason) %>%
  arrange(desc(n)) %>%  # arranges in descending order
  adorn_pct_formatting(digit = 2) # rounds to 2 decimal digits
```

```{r}
#| label: q1.b
#////////////////////////////////////////////////////////////////////////////
### Make one bar chart showing the number of discharges by services_d (type of treatment service). Give it an informative title and axis labels.

##loads packages if not done so already
library(tidyverse)
library(janitor) 
library(naniar) # tools for exploring and visualizing missing data
library(plotly) # makes finished ggplot charts interactive

###MY VERSION
freq_p <- ggplot(data = teds_d, aes(x = fct_infreq(services_d))) + #fct_infreq sorts by largest to smallest
  geom_bar() +
  coord_flip() + # flips cooordinates; can use y = within aes() instead too
  labs(x = "Type of treatment service")
ggplotly(freq_p)

###EASY/SIMPLE VERSION
ggplot(teds_d, aes(y = fct_infreq(services_d))) +  #changed "x" to "y" for different bar orientation
  geom_bar()

###AMBER VERSION
ggplot(teds_d, aes(y = fct_infreq(services_d))) + #changed y to servies instead of x
  geom_bar() +
  scale_x_continuous(labels = scales::label_comma()) + #adds commands to values on axis
  labs(title = "Treatment Discharges by Type of Service, 2024 TEDS-D",
       x = "Number of discharges",
       y = "Type of treatment servive") +
  theme_minimal()

### In a sentence or two: what does the reason value "Transferred to another treatment program or facility" mean for our completion percentages? If a state transfers a lot of patients, would its completion rate look higher or lower than a state that transfers very few?

# overall % of each is higher due to less overall patients within record
# if they stay in data set, doesn't change much
```

**Answer to (a): refer to label: q1.a**

**Answer to (b):** (produce a bar chart in the code chunk above)

**Answer to (c):** "Transferred to another treatment program or facility" may or may not mean something for the completion percentages, depending on how the "Transferred to (...)" designation is handled.

|  |  |  |
|----|----|----|
| **is treatment complete** | **record updated locally** | **result** |
| Y | Y | increase in local completion |
| Y | N | no change in local, recorded still as transferred |
| N | Y | no change in local, recorded still as transferred |
| N | N | no change in personal, recorded still as transferred |

**If a state transfers a lot of patients, would its completion rate look higher or lower than a state that transfers very few?**

For a state with many transfer patients, the completion rate (%) would look higher when the majority of patients that transferred have not completed treatment, and lower when the majority of patients that transferred actually completed treatment.

## Question 2: Mapping no use at discharge

Make **one [*interactive*]{.underline} map** with `ggiraph` showing, for each state, the percentage of [completed treatment]{.underline} episodes that ended with [no use at discharge]{.underline}.

In other words: among the episodes where `reason` is "Treatment completed," what share had `freq1_d` equal to "no use"?

Work through this in bite-sized steps, like we did in Part 2. You will need to filter, group, summarize, build an index, join, and plot. Write one step at a time and look at the result before moving on.

```{r}
#####FOR QUESTION 2
# only install these packages if you don't already have them!
install.packages("tidyverse")
install.packages("janitor")
install.packages("sf")
install.packages("remotes") # used to install packages from github
remotes::install_github("UrbanInstitute/urbnmapr") # use this to install urbnmapr
install.packages("naniar")
install.packages("ggiraph")
```

```{r, message=FALSE, warning=FALSE}
#loading packages
library(tidyverse)
library(janitor) 
library(sf) # "simple features" for spatial data
library(urbnmapr) # access to US state/country shapefiles for mapping
library(naniar) # tools for exploring and visualizing missing data
library(ggiraph) # makes ggplot charts interactive
options(scipen = 99) # turns off scientific notation
```

```{r}
#| label: q2 start

# selecting data
my_teds_d <- teds_d %>%
  clean_names() %>% # janitor
  select(stfips, freq1_d, reason) # only focusing on these this week

head(my_teds_d)
```

```{r}
#missing-data-w-naniar
miss_var_summary(my_teds_d)

gg_miss_var(my_teds_d)

my_teds_d %>%
  slice_sample(n = 500) %>% # from dplyr
  vis_miss() # from naniar

teds_d %>% 
  tabyl(freq1_d) %>% # from janitor package
#NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked
```

```{r}
#filtering by treatment completed AND removing entries which have "unknown" listed under the freq1_d column
filtered_my_teds_d <- my_teds_d %>%
  dplyr::filter(
    reason == "Treatment completed",#, # keeping "Treatment completed" entries from reason column
    freq1_d != "unknown" # keep non-'unknown' entries from freq1_d column
    #culmen_length_mm >= 39,
    #island %in% c("Dream", "Torgersen")
  )
filtered_my_teds_d

#####CHECKING FOR ANY REMAINING "UNKNOWN" ENTIRES THAT SHOULD BE REMOVED
sum(filtered_my_teds_d$freq1_d == "unknown", na.rm = TRUE)
###METHOD TO VISUAL CHECK
filtered_my_teds_d %>% 
  tabyl(freq1_d) %>% # from janitor package
  #NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked
```

```{r}
#| label: percentage per state
#grouping the treatment completed cleaned dataset by state

percent_completed_by_state <- filtered_my_teds_d |> #takes teds dataset, each row is a specific record (left for a various reason)
  group_by(stfips) |> # group by stfips code
  summarize(total_cases = n(), # n() is to count rows/TOTAL ROWS
            no_use_cases = sum(freq1_d == "no use", na.rm = TRUE)) |>  # na.rm removes na values if they exist
  mutate(percentage_completed = (no_use_cases / total_cases) * 100)

summary(percent_completed_by_state)
percent_completed_by_state

##### AMBER LOOK HERE
#### output displays that for stfips #4, its no_use_cases has 0 value
```

```{r}
#| label: DOESN'T WORK; checking individual state for count "no use"

###counting total 'stfips == 4' entries
testStfips4 <- filtered_my_teds_d %>%
  #  filter(stfips == 4) %>%  # Checks for specific variable in Column 1
  dplyr::count(stfips == 4)                      # Counts a different variable in Column 2
print(testStfips4)

###counting "no use" entries of 'stfips == 4'
testStfips4 <- filtered_my_teds_d %>%
  filter(stfips == 4) %>%  # Checks for specific variable in Column 1
  count("no use")                      # Counts a different variable in Column 2
print(testStfips4)

######DOESN'T WORK; IM RETARDED
######can't put conditional check in count();
######NO IDEA WHAT THE ABOVE CODE IS EVEN CHECLKING FOR
```

```{r}
#| label: WORKS; checking individual state for count "no use"

# checking values from above since some completed cases are 0

# multi-conditional search using summarize, NO FILTER()
testStfips4_2 <- filtered_my_teds_d %>%
  summarize(
    matching_count = sum(stfips == 4 & freq1_d == "no use")  #changed "status" to "freq1_d"
  )
#print(testStfips4_2) # PRINTS IN CONSOLE
testStfips4_2

# sequential searches using filter first
testStfips4_2 <- filtered_my_teds_d %>%
  filter(stfips == 4)
sum(testStfips4_2$freq1_d == "no use", na.rm = TRUE)

# nested sequential searches using filter + summarise
testStfips4_2 <-filtered_my_teds_d %>%
  filter(stfips == 4) %>%
  summarise(count = sum(freq1_d == "no use", na.rm = TRUE))
testStfips4_2

#####WORKS NOW, count is correct; stfips 4 does not have any explicitly defined cases with "no use"
```

```{r}
#| label: check-percentages

#check math of any%; average of average between states
mean(percent_completed_by_state$percentage_completed)

#### fraction of completed over total
sum(percent_completed_by_state$no_use_cases) /
  sum(percent_completed_by_state$total_cases) * 100

summary(percent_completed_by_state$percentage_completed)

sum(percent_completed_by_state$total_cases) == nrow(filtered_my_teds_d)

```

#####GETTING MAP DATA

```{r}
#| message: false
#| warning: false
#####GETTING MAP DATA

states_map <- get_urbn_map(map = "states", sf = TRUE)

head(states_map, 10)
```

```{r}

#####preparing index
# copies stfips data to create new column in percent_completed_by_state called 
# state_fips, forcing values to be two digits (start with zero)

percent_completed_by_state <- percent_completed_by_state |> # take index from percent_completed_by_state
  mutate(state_fips = sprintf("%02d", stfips)) # sprintf = string print formatted
#%02d sets to 2 digits

head(percent_completed_by_state)

#converted to TEXT VALUES, or the 0 in front would be dropped

```

```{r}
#| message: false
#| warning: false
#| 
#####stitching for index
mapping_data <- full_join(percent_completed_by_state,
                          states_map,
                          by = "state_fips") # remember the index we prepared above?
```

```{r}
##### comparing row counts

nrow(percent_completed_by_state)
nrow(states_map)
nrow(mapping_data)
```

#####PLOTTING MAP

```{r}
##### interactive map + cleaned hover tooltip

completed_treatment_map_interactive <- ggplot(mapping_data) +
  geom_sf_interactive(mapping = aes(geometry = geometry,
                                    fill = percentage_completed,
                                    tooltip = paste(
                                      "State:", state_abbv,
                                      "<br>State FIPS:", stfips, 
                                      "<br>Completed:", round(percentage_completed, 2), "%")),
                      color = "#ffffff",
                      linewidth = 0.25) + # changes border smaller for easier visibility
  labs(fill = "% of Completed Treatment Episodes ending with verified no use at discharge",
       title = "Completed Treatment Episodes with no use at discharge by State",
       subtitle = "TEDS-D 2024 (SAMHSA)") +
  scale_fill_viridis_c(option = "mako", direction = -1) + #direction flips direction of colors; changes higher = more saturation
    #####fill data is a continuous sequence of numbers (like temperature, heights, or counts), replace scale_fill_viridis_d() with the continuous version, scale_fill_viridis_c()
  coord_sf(datum = NA) + # turns off lat/long grid
  theme_minimal() +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.text = element_text(size = 4),
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4))


# Use `girafe` to render the interactive plot
girafe(ggobj = completed_treatment_map_interactive)
```

DENOMINATOR I USED: I chose to not include the "unknown" values of freq1_d as that is what is considered as the best practice to follow when dealing with the subject matter of standard public health reporting and data analytics.\
Including uncertain (in this case, unknown) data in the total count would artificially deflate the calculated percentage, which would be a misrepresentation of the accurately recorded data.

### The denominator decision

About a third of the `freq1_d` values in this dataset are "unknown." That means you have to decide what your denominator is, and there is more than one defensible answer:

- **All** completed episodes, counting "unknown" in the denominator, or

- Only completed episodes where frequency of use was **actually recorded**.

Pick one. **Below your map, write one sentence saying which denominator you used and why.** Either choice can earn full credit. Missing description earns no points for this portion.

### Your map must:

- be interactive (rendered with `girafe()`)

- show a tooltip with the state name and the percentage, rounded to two decimal places

- have a title and a legend label that say what is being shown

- handle states with no data honestly (remember Friday in class)

> **Full credit:** one interactive map meeting the four requirements above, plus your denominator sentence. It does not have to be beautiful, only correct and readable.

```{r}
#| label: q2.1

```

**Denominator I used, and why:**

## Question 3: Service type and length of stay

Treatment outcomes can vary depending on the type of service a patient receives and how long they stay in treatment.

Investigate how **completion rates** differ across **service type** (`services_d`) and **length of stay** (`los`).

- Create **two** visualizations: at least one focused on service type, and at least one focused on length of stay. These do not need to be maps, and they do not need to be interactive. Pick whatever chart type makes the pattern easiest to see.

- Then write a short summary (3-ish sentences, full grammatical sentences) explaining the main patterns you observe.

**One warning:** look up `los` in the codebook *before* you calculate anything with it. Code that runs without an error can still give you a meaningless answer.

> **Full credit:** two labeled visualizations that actually show the comparison you say they show, plus a summary that describes a real pattern in your own output. You are not graded on finding the "right" pattern. You are graded on whether your summary matches your charts.

```{r}
#| label: q3.1
#####completion rates vs service type (services_d)
#bar

# selecting data
my_teds_d_services_d <- teds_d %>%
  clean_names() %>% # janitor
  select(services_d, reason)

head(my_teds_d_services_d)
```

```{r}
#missing-data-w-naniar
miss_var_summary(my_teds_d_services_d)

gg_miss_var(my_teds_d_services_d)

my_teds_d_services_d %>%
  slice_sample(n = 500) %>% # from dplyr
  vis_miss() # from naniar

teds_d %>% 
  tabyl(services_d) %>% # from janitor package
  #NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked
```

```{r}
#filtering by treatment completed
filtered_my_teds_d_services_d <- my_teds_d_services_d %>%
  dplyr::filter(
    reason == "Treatment completed",#, # keeping "Treatment completed" entries from reason column
    #culmen_length_mm >= 39,
    #island %in% c("Dream", "Torgersen")
  )
##############checking to ensure all entries are "Treatment completed"
#####IGNORE THIS:CHECKING FOR ANY REMAINING "UNKNOWN" ENTIRES THAT SHOULD BE REMOVED; NEED TO MODIFY IF USED
#sum(filtered_my_teds_d_services_d$freq1_d == "unknown", na.rm = TRUE)
###METHOD TO VISUAL CHECK
filtered_my_teds_d_services_d %>% 
  tabyl(reason) %>% # from janitor package
  #NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked

filtered_my_teds_d_services_d

#sum() will print out result in console
sum(filtered_my_teds_d_services_d$reason == "Treatment completed", na.rm = TRUE)
#####CHECK "WORKS; checking individual state for count "no use"" for other methods of checking
```

```{r}
### Using tabyl(), show the breakdown of services_d (services_d for discharge). Which services_d is most common, and what percentage of all discharges does it account for?
class(filtered_my_teds_d_services_d$services_d)
#look at this variable; check class
summary(filtered_my_teds_d_services_d$services_d)
#summary of data frame

filtered_my_teds_d_services_d <- filtered_my_teds_d_services_d %>%
  mutate(services_d = recode(as.character(services_d),
                             "1" = "Detox, 24-hour, hospital inpatient",
                             "2" = "Detox, 24-hour, free-standing residential",
                             "3" = "Rehab/residential, hospital (non-detox)",
                             "4" = "Rehab/residential, short term (30 days or fewer)",
                             "5" = "Rehab/residential, long term (more than 30 days)",
                             "6" = "Ambulatory, intensive outpatient",
                             "7" = "Ambulatory, non-intensive outpatient",
                             "8" = "Ambulatory, detoxification",
                             .missing = "unknown"),
         services_d = factor(services_d))
# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

# organizes relevant info
filtered_my_teds_d_services_d %>% 
  tabyl(services_d) %>%
  arrange(desc(n)) %>%  # arranges in descending order
  adorn_pct_formatting(digit = 2) # rounds to 2 decimal digits
```

```{r}
##loads packages if not done so already
library(tidyverse)
library(janitor) 
library(naniar) # tools for exploring and visualizing missing data
library(plotly) # makes finished ggplot charts interactive
```

```{r}
###MY VERSION BAR CHART
plot_filtered_my_teds_d_services_d <- ggplot(data = filtered_my_teds_d_services_d, aes(x = fct_infreq(services_d))) + #fct_infreq(variable) sorts by largest to smallest  
    ######REORDER DOESN'T WORK; ASK WHY
  geom_bar() +
  coord_flip() + # flips cooordinates; can use y = within aes() instead too
  labs(title = "Completed Treatment Discharges by Type of Service, 2024 TEDS-D",
       x = "Number of discharges",
       y = "Type of treatment service") +
  theme(plot.title = element_text(hjust = 0.5)) + #doesn't shift title to center??
  theme_minimal()
ggplotly(plot_filtered_my_teds_d_services_d)
```

```{r}

```

```{r}
#| label: q3.2
#####completion rates vs length of stay (los)
#histogram

#####completion rates vs service type (services_d)
#bar

# selecting data
my_teds_d_los <- teds_d %>%
  clean_names() %>% # janitor
  select(stfips, los, reason)

head(my_teds_d_los)
```

```{r}
#missing-data-w-naniar
miss_var_summary(my_teds_d_los)

gg_miss_var(my_teds_d_los)

my_teds_d_los %>%
  slice_sample(n = 500) %>% # from dplyr
  vis_miss() # from naniar

teds_d %>% 
  tabyl(los) %>% # from janitor package
  #NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked
```

```{r}
#filtering by treatment completed
filtered_my_teds_d_los <- my_teds_d_los %>%
  dplyr::filter(
    reason == "Treatment completed",#, # keeping "Treatment completed" entries from reason column
    #culmen_length_mm >= 39,
    #island %in% c("Dream", "Torgersen")
  )
##############checking to ensure all entries are "Treatment completed"
#####IGNORE THIS:CHECKING FOR ANY REMAINING "UNKNOWN" ENTIRES THAT SHOULD BE REMOVED; NEED TO MODIFY IF USED
#sum(filtered_my_teds_d_services_d$freq1_d == "unknown", na.rm = TRUE)
###METHOD TO VISUAL CHECK
filtered_my_teds_d_los %>% 
  tabyl(reason) %>% # from janitor package
  #NO MISSING DATA; CAN MOVE ON
  adorn_totals("row") # shows total of variables being checked

filtered_my_teds_d_los

#sum() will print out result in console
sum(filtered_my_teds_d_los$reason == "Treatment completed", na.rm = TRUE)
#####CHECK "WORKS; checking individual state for count "no use"" for other methods of checking
```

```{r}
### Using tabyl(), show the breakdown of services_d (services_d for discharge). Which services_d is most common, and what percentage of all discharges does it account for?
class(filtered_my_teds_d_los$los)
#look at this variable; check class
summary(filtered_my_teds_d_los$los)
#summary of data frame

filtered_my_teds_d_los2 <- filtered_my_teds_d_los %>%
  mutate(value_los = los,
         value_los = recode(as.character(value_los),
                            "1" = "1" ,
                            "2" = "2" ,
                            "3" = "3" ,
                            "4" = "4" ,
                            "5" = "5" ,
                            "6" = "6" ,
                            "7" = "7" ,
                            "8" = "8" ,
                            "9" = "9" ,
                            "10" = "10",
                            "11" = "11",
                            "12" = "12",
                            "13" = "13",
                            "14" = "14",
                            "15" = "15",
                            "16" = "16",
                            "17" = "17",
                            "18" = "18",
                            "19" = "19",
                            "20" = "20",
                            "21" = "21",
                            "22" = "22",
                            "23" = "23",
                            "24" = "24",
                            "25" = "25",
                            "26" = "26",
                            "27" = "27",
                            "28" = "28",
                            "29" = "29",
                            "30" = "30",
                            "31" = "45",
                            "32" = "60",
                            "33" = "90",
                            "34" = "120",
                            "35" = "180",
                            "36" = "365",
                            "37" = "366",
                            .missing = "unknown"),
         value_los = factor(value_los),
         label_los = los,
         label_los = recode(as.character(label_los),
                            "1" = "1" ,
                            "2" = "2" ,
                            "3" = "3" ,
                            "4" = "4" ,
                            "5" = "5" ,
                            "6" = "6" ,
                            "7" = "7" ,
                            "8" = "8" ,
                            "9" = "9" ,
                            "10" = "10",
                            "11" = "11",
                            "12" = "12",
                            "13" = "13",
                            "14" = "14",
                            "15" = "15",
                            "16" = "16",
                            "17" = "17",
                            "18" = "18",
                            "19" = "19",
                            "20" = "20",
                            "21" = "21",
                            "22" = "22",
                            "23" = "23",
                            "24" = "24",
                            "25" = "25",
                            "26" = "26",
                            "27" = "27",
                            "28" = "28",
                            "29" = "29",
                            "30" = "30",
                            "31" = "31 ~ 45",
                            "32" = "46 ~ 60",
                            "33" = "61 ~ 90",
                            "34" = "91 ~ 120",
                            "35" = "121 ~ 180",
                            "36" = "181 ~ 365",
                            "37" = "365+",
                            .missing = "unknown"),
         label_los = factor(label_los),
         los = sprintf("%03d", los))
# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

# organizes relevant info
filtered_my_teds_d_los2 %>% 
  tabyl(los) %>%
  arrange((los))# %>%  # arranges in ascending order
#  #adorn_pct_formatting(digit = 2) # rounds to 2 decimal digits
```

```{r}
##loads packages if not done so already
library(tidyverse)
library(janitor) 
library(naniar) # tools for exploring and visualizing missing data
library(plotly) # makes finished ggplot charts interactive
```

```{r}
###MY VERSION BAR CHART
plot_filtered_my_teds_d_services_d <- ggplot(data = filtered_my_teds_d_services_d, aes(x = fct_infreq(services_d))) + #fct_infreq(variable) sorts by largest to smallest  
  ######REORDER DOESN'T WORK; ASK WHY
  geom_bar() +
  coord_flip() + # flips cooordinates; can use y = within aes() instead too
  labs(title = "Completed Treatment Discharges by Type of Service, 2024 TEDS-D",
       x = "Number of discharges",
       y = "Type of treatment service") +
  theme(plot.title = element_text(hjust = 0.5)) + #doesn't shift title to center??
  theme_minimal()
ggplotly(plot_filtered_my_teds_d_services_d)
```

```{r}
library(ggplot2)

# Plot using density
ggplot(filtered_my_teds_d_los2, aes(x = los)) +
  geom_histogram(aes(y = after_stat(density)), 
                 breaks = c(0, 1, 2, 3, 4, 5, 10), 
                 fill = "steelblue", 
                 color = "white") +
  labs(title = "Duration Histogram",
       x = "Years",
       y = "Density") +
  theme_minimal()
```

```{r}
library(ggplot2)

# Create a data frame
df <- data.frame(durations = c(1, 1.5, 2, 3, 4, 5, 6, 7, 8, 9, 10), density = c(1,2,3,4,5,6,7,8,9,10,11))

# Plot using density
ggplot(df, aes(x = durations)) +
  geom_histogram(aes(y = after_stat(density)), 
                 breaks = c(0, 1, 2, 3, 4, 5, 10), 
                 fill = "steelblue", 
                 color = "white") +
  labs(title = "Duration Histogram",
       x = "Years",
       y = "Density") +
  theme_minimal()
```

**Summary:**

## Extra Credit (optional, up to 2 pts)

#### Again, this is meant to *stretch* you, not *stress* you!

Make one visualization that shows service type and length of stay **together**–that is, how completion rate varies across both at once. Faceting is one way, but there are others.

Then, in a sentence, say what this combined view shows that your two separate charts in Question 3 could not.

```{r}
#| label: extra-credit

```

**What the combined view adds:**

------------------------------------------------------------------------

## Push to GitHub AND submit link on Canvas!

Assignment is due **Friday 10/2, 11:59pm**.
