---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### LOAD LIBRARY

```{r}
library(ggplot2)

```

### Some scripting

```{r}
# this is a comment
quiz1 <- 88 # this is how to assign a value to a name
quiz2 <- 94
quiz3 <- 79

quiz1

# average
(quiz1 + quiz2 + quiz3)/3

mean(quiz1 + quiz2 + quiz3) # NOT CORRECT; ADDS THEM
mean(quiz1, quiz2, quiz3) # NOT CORRECT; TAKES ONLY MEAN OF FIRST ARGUMENT
mean(c(quiz1, quiz2, quiz3)) # CORRECTLY TAKES MEAN; C IS A LIST

myList <- c(quiz1, quiz2, quiz3) # assigns list to variable "myList"

mean(myList)

average <- mean(myList)

quiz3 <- 99 # changes value of variable "quiz3" to 99, but has not updated "average" OR "myList"

myList <- c(quiz1, quiz2, quiz3) # UPDATING ALL VARIABLES/FUNCTIONS
average <- mean(myList)
average
round(average, 2)

```

### More scores!

```{r}
myScores <- c(88, 94, 79, 91, 85)
# myScores ; commented due to breaking

mean(myScores)
min(myScores)
max(myScores)
length(myScores)
summary(myScores)

myStudents <- c("A", "B", "C", "D", "E")
# myStudents; commented due breaking

# bar plot
barplot(myScores, names.arg = myStudents, ylim = c(0, 100), main = "TITLE HERE", xlab = "pupils", y = "Test Scores")

```

## Introducing `ggplot2`

#### Make a dataframe

```{r}

grades_df <- data.frame(student = myStudents, score = myScores)
# make list of students using myStudents, and scores using myScores
grades_df

summary(grades_df)

nrow(grades_df) # numb of rows
ncol(grades_df)
dim(grades_df) # df dimensions
```

#### Try ggplot

```{r}

ggplot(data = grades_df, aes(student, y = score)) # foundation

ggplot(data = grades_df, aes(x = reorder(student, -score), y = score, group = 1)) + # needed to add group = 1 for geom_line to tell it that it is part of the same group
  # aes = aesthetic
  # for reorder function, can just leave it as student to not reorder
  geom_col(fill = "steelblue") +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "BEST GRAPH EVER", # labels
       subtitle = "some quiz scores",
       x = "the students",
       y = "what they scored on their quiz")
```

#### Mutate dataframe

```{r}

library(tidyverse)

grades_df <- grades_df |> # makes new graph, first variable can be old named graph to overwrite
  mutate(study_hours = c(2, 5, 1, 4, 3))
```

#### Make scatterplot

```{r}

ggplot(data = grades_df, aes(x = study_hours, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# can't use fill (use color instead), don't use reorder
```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

1 + 1 + 1

quiz1 + quiz2


1 + 1 + 1 - 3

0.1 + 0.1 + 0.1 - 0.3
# why is this not 0

options(scipen = 999) # turns off scientific  notatiion

print(0.1, 20) # prints 0.1 value with 20 decimal
print(0.3, 20) # also doesn't print properly
print(0.5, 20) # it is acutally 0.5 in this case

# {python} if used, then will still give same value

### ANSWER: values are stored in base-2 instead of base-10 AKA binary; can't be represented accurately, so they are slightly off.
# solution: use float

# source for answer
# https://www.reddit.com/r/learnpython/comments/ly5jv4/whats_different_between_decimal_and_floating/
# https://docs.python.org/3/library/decimal.html
```

What did you observe?

> Answer: Solutions were close but not exact values.

Why do you think this is happening? (You can look online for help)

> Answer: Values are stored in base-2 (binary) instead of base-10; can't be represented accurately, so they are slightly off.
>
> Solution: use float instead, or different type.
