---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### LOAD LIBRARY

```{r}
library(ggplot2)

```

### Some scripting

```{r}
# this is a comment
quiz1 <- 88 # this is how to assign a value to a name
quiz2 <- 94
quiz3 <- 79

quiz1

# average
(quiz1 + quiz2 + quiz3)/3

mean(quiz1 + quiz2 + quiz3) # NOT CORRECT; ADDS THEM
mean(quiz1, quiz2, quiz3) # NOT CORRECT; TAKES ONLY MEAN OF FIRST ARGUMENT
mean(c(quiz1, quiz2, quiz3)) # CORRECTLY TAKES MEAN; C IS A LIST

myList <- c(quiz1, quiz2, quiz3) # assigns list to variable "myList"

mean(myList)

average <- mean(myList)

quiz3 <- 99 # changes value of variable "quiz3" to 99, but has not updated "average" OR "myList"

myList <- c(quiz1, quiz2, quiz3) # UPDATING ALL VARIABLES/FUNCTIONS
average <- mean(myList)
average
round(average, 2)

```

### More scores!

```{r}
myScores <- c(88, 94, 79, 91, 85)
myScores

mean(myScores)
min(myScores)
max(myScores)
length(myScores)
summary(myScores)

myStudents <- c("A", "B", "C", "D", "E")
myStudents

# bar plot
barplot(myScores, names.arg = myStudents)
```

## Introducing `ggplot2`

```{r}

```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
