---
title: "TEDS-D Treatment Mapping"
author: "Amber Camp" 
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Load Packages

If you don't have the below packages, run the relevant lines here:

```{r}
#| label: install-packages

# we'll use all of these this week, but selectively load a few today

###don't need to use these anymore once done
# install.packages("sf")
# install.packages("remotes") # used to install packages from github
# remotes::install_github("UrbanInstitute/urbnmapr")
# install.packages("naniar")
# install.packages("ggiraph")
```

Load packages (all of them).

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
library(janitor) 
library(naniar) # tools for exploring and visualizing missing data
library(plotly) # makes finished ggplot charts interactive
options(scipen = 99) # turns off scientific notation
```

## Today's Data

The data we will analyze today is the 2024 Treatment Episode Data Set-Discharges (TEDS-D) Dataset from the Substance Abuse and Mental Health Services Administration (SAMHSA). The metadata can be found [here](https://www.samhsa.gov/data/system/files/media-puf-file/TEDS-D-2024-DS0001-info-codebook_v1_0.pdf) CODE BOOK HERE

This is a huge dataset! Almost 1.3 million discharge records and 76 variables. SAMHSA distributes it as an `.rdata` file, which we read in with `load()`.

`load()` works differently from functions like `read_csv()`. It doesn't hand the data back to you. Instead, it recreates the saved object in your environment under whatever name it was *originally saved with*, and returns that name as text. We capture the name so we can grab the object with `get()`.

```{r}
#| label: load-data

loaded <- load("data/tedsd_puf_2024.rdata")

loaded # what was the object called?

teds_d <- get(loaded)
rm(list = loaded) # remove the extra copy from your environment

dim(teds_d)
```

### Keep only what we need

The full dataset takes up nearly a gigabyte of memory. Big datasets are common in the real world, and a good habit is to keep only the columns you actually need as early as possible. That makes everything after this step faster.

```{r}
#| label: select-data

teds_d <- teds_d %>%
  clean_names() %>% # janitor
  select(stfips, freq1_d, services_d, los, reason) # only focusing on these this week

head(teds_d)
```

### Columns for today's class

- `freq1_d` = Frequency of use at discharge

- `stfips` = State FIPS code (FIPS = *Federal Information Processing Standard)*

- `services_d` = Treatment Service

- `los` = Length of Stay

- `reason` = Reason for Discharge

### NA Analysis

What data is missing? Examine and visualize below with `naniar`

```{r}
#| label: missing-data-w-naniar

miss_var_summary(teds_d)

gg_miss_var(teds_d)

teds_d %>%
  slice_sample(n = 500) %>% # from dplyr
  vis_miss() # from naniar
```

There actually is more missing data than the above would suggest... Let's view a table summary.

```{r}
#| label: freq1_d-table

teds_d %>% 
  tabyl(freq1_d) # from janitor package
```

How does the documentation label missing data? Replace values as needed.

With a dataset this size, we replace the missing codes one column at a time with `na_if()`. A shortcut like `teds_d[teds_d == -9] <- NA` compares every cell at once and can freeze your laptop on 1.3 million rows.

```{r}
#| label: replace-missing-codes

teds_d <- teds_d %>%
  mutate(across(everything(), ~ na_if(.x, -9)))
    #.x is where we are atm, ~ is what to do

# read like this:
# "Take teds_d, and then mutate it: across every column, replace -9 with NA. 
# Save the result back as teds_d."

# Note: the ~ says "here's what to do to each column," 
# and .x means "the column we're on right now"
```

Inspect once again.

```{r}
#| label: inspect-missing-values

miss_var_summary(teds_d)

teds_d %>% 
  tabyl(freq1_d)
```

Now check `reason`. Does it have any missing values in the 2024 data? Does that match what the codebook says?

```{r}
#| label: inspect-reason-code

miss_var_summary(teds_d)

teds_d %>% 
  tabyl(reason)

```

### Variable Re-coding

Many of the variables are currently in a format that is not helpful for our analyses.

#### Frequency of Use at Discharge

Start by inspecting this variable. Next, look up the Frequency of Use at Discharge variable in the codebook (`freq1_d`)

```{r}
#| label: variable-recoding-freq1_d

class(teds_d$freq1_d)
  #look at this variable; check class
summary(teds_d$freq1_d)
  #summary of data frame

teds_d <- teds_d %>%
  mutate(freq1_d = recode(as.character(freq1_d),
                          "1" = "no use",
                          "2" = "some use",
                          "3" = "daily use",
                          .missing = "unknown"),
         freq1_d = factor(freq1_d))

# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

teds_d %>% 
  tabyl(freq1_d) 
```

#### Services

Apply the relevant recoding for `services_d`

```{r}
#| label: variable-recoding-services_d

class(teds_d$services_d)
#look at this variable; check class
summary(teds_d$services_d)
#summary of data frame

teds_d <- teds_d %>%
  mutate(services_d = recode(as.character(services_d),
                          "1" = "Detox, 24-hour, hospital inpatient",
                          "2" = "Detox, 24-hour, hospital inpatient",
                          "3" = "Rehab/residential, hospital (non-detox)",
                          "4" = "Rehab/residential, short term (30 days or fewer)",
                          "5" = "Rehab/residential, long term (more than 30 days)",
                          "6" = "Ambulatory, intensive outpatient",
                          "7" = "Ambulatory, non-intensive outpatient",
                          "8" = "Ambulatory, detoxification",
                          .missing = "unknown"),
         services_d = factor(services_d))

# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

teds_d %>% 
  tabyl(services_d) 


```

#### Reason

Apply the relevant recoding for the Reason for Discharge variable.

```{r}
#| label: variable-recoding-reason

class(teds_d$reason)
#look at this variable; check class
summary(teds_d$reason)
#summary of data frame

teds_d <- teds_d %>%
  mutate(reason = recode(as.character(reason),
                          "1" = "Treatment completed",
                          "2" = "Dropped out of treatment",
                          "3" = "Terminated by facility",
                          "4" = "Transferred to another treatment program or facility",
                          "5" = "Incarcerated",
                          "6" = "Death",
                          "7" = "Other",
                          .missing = "unknown"),
         reason = factor(reason))

# "Turn the codes into text, swap each code for its label (and call missing ones 'unknown'),
# then treat the result as a category (factor)."

teds_d %>% 
  tabyl(reason) 

```

## Quick Plotting Exercise

Use `plotly` to create three basic visualizations using `ggplot2` and the variables you recoded above (`freq1_d`, `services_d`, and `reason`). Bar plots are a good idea :)

Tips:

- use `fct_infreq()` around your `x` variable to organize in descending order of count

- try using `coord_flip()` for a horizontal view

- don't forget an axis label!

```{r}
#| label: plots

freq_p <- ggplot(data = teds_d, aes(x = fct_infreq(freq1_d))) +
  geom_bar() +
  coord_flip() + # not required
  labs(x = "Frequency of Use at Discharge")

ggplotly(freq_p)

###
services_p <- ggplot(data = teds_d, aes(x = fct_infreq(services_d))) +
  geom_bar() +
  coord_flip() + # not required
  labs(x = "Treatment Service")

ggplotly(services_p)

###
reason_p <- ggplot(data = teds_d, aes(x = fct_infreq(reason))) +
  geom_bar() +
  coord_flip() + # not required
  labs(x = "Reason for discharge")

ggplotly(reason_p)
```

## Do before Wednesday's Class

### Save your new dataframe

You've put a lot of work into cleaning up this dataframe! Once you're done recoding everything, save it to a new `.rds` file so that you can simply read it in on Wednesday! (`.rds` files store a single R object, and they keep your factors intact.)

Have this done before class on Wednesday.

```{r}
#| label: save-dataframe

saveRDS(teds_d, "data/teds_d_clean.rds")
#
```
